Lattice Boltzmann simulations of magnetically responsive\\ ellipsoidal particle stabilized emulsions

### 2025-03-24 chapters/LBM_technique.tex
Changes 1. Reworded introduction and lit review and addressed grammar and vocab issues 2. Fixed some referencing problems in the document 3. Fleshed out introduction in paper 2 4. Added sections for results in paper 3
Changes 1. Added conclusion in final remarks 2. Added hysteresis curve for prolate particles stabilized bijels in aim2 section 3. Added description of synthesis techniques in lit review
Changes 1. Refined text in most sections 2. Resized figures for consistency of figure size 3. Removed shear banding analysis and added how shear stress is calculated
Changes 1. Modified and added

\label{section:methods}

@@

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale = 0.7]{figures/methods/d3q19_lattice.jpg}
    \caption{D3Q19 lattice demonstrating the rest, nearest and next nearest direction that correspond to the 19 
    directions $(i)$ of the lattice with lattice velocity $c_{i}$. \cite{schmieschek_lb3d_2017} Reproduced from 
    Schmieschek et al. under the Creative Commons license.}
    \label{fig:d3q19_lattice}
\end{figure}

@@

>>>>

\section{Methods} \label{section:methods}

@@

\section{Lattice Boltzmann Method}

The Lattice Boltzmann Method (LBM) is an evolution of preceding lattice gas automata techniques, which is a discretization of the Boltzmann equation of motion for 
molecules. This means that unlike traditional CFD techniques such as FDM, VOF or level set methods, LBM is a psuedo-molecule method that tracks the evolution of a 
particle distribution function within grid cells evolved through a discretized Boltzmann equation of motion. Macroscopic variables such as fluid density and 
velocity are recovered from the particle distributions through appropriate moment integration, and the Navier Stokes equation at the incompressible limit can be 
obtained through a Chapman-Enskog expansion of the LBM. The LBM has become an attractive tool for meso-scale CFD simulations due to its ease of algorithm 
implementation, highly parallelizable nature and ease of boundary condition implementation, allowing coupling to other physically relevant systems such as 
particles with varieties of potentials, external fields and deformable bodies.

The particle distribution function described earlier is advected on a pre-constructed lattice stencil, commonly denoted as DnQm where n and m represent the 
number of dimensions and directions in the stencil. Common stencils include the D1Q5, D2Q9 and D3Q19 stencil, all of which recover mass and momentum conservation. 
For energy conservation, a higher order stencil such as D3Q27 is necessary. The D2Q9 and D3Q19 stencils have 8 and 18 populations respectively that include 
connections to nearest and next nearest neighbour points, in addition to a central rest point.

The LBM is composed of a collision and advection step. In the advection step, the populations at each grid point are propagated to adjacent points in accordance 
with the chosen stencil. During the collision step, the particle population distribution is relaxed towards an equilibrium with a collision operator, at a 
specified relaxation rate. The collision operator can have multiple forms based on how many relaxation rates are used although the most common variant is the 
Single Relaxation Time (SRT) collision operator, more commonly known as the Bhatnagar-Gross-Krook (BGK) collision operator. Owing to its stability in low Mach 
and Reynolds numbers and simplicity of implementation, the BGK operator is often used in common particle laden flow and soft matter scenarios. 
\cite{bhatnagar_model_1954} These limitations are present owing to the implicit link between the fluid properties and the relaxation rate necessitating 
relaxation rates above 0.5, and to ensure fluid incompressibility from an equation that intrinsically simulates a compressible fluid. To get over these 
limitations, Two Relaxation Time (TRT) and Multiple Relaxation Time (MRT) operators also exist, expanding the possible application of the LBM to visco-elastic 
flows and implementation of fluctuating hydrodynamics in the LBM. \cite{liu_simulation_2023, adhikari_fluctuating_2005}

Four primary techniques to model multicomponent or multiphase systems exist in the LBM literature. A quick review of the Shan-Chen or 
interparticle potential model  will be provided here from the perspective of a binary mixture. Additionally, the names and descriptions of 
other techniques will be described as well. The interparticle potential model or Shan-Chen model that adds a non-local density dependent force 
between two species, effectively modelling non-ideal mixing and recovering the Cahn-Hilliard equation. To alleviate the standard Shan-Chen implementations 
weaknesses of not being thermodynamically consistent and reducing the existence of spurious velocities,

In addition to the Shan-Chen model, the color gradient model, free energy model and interface tracking technique all 
allow for simulations of various types of multiphase and/or multicomponent flows. The color gradient model proposed by Rothman and Keller and 
implemented by Gunstensen et al relies upon modelling two particle distributions. These represent a binary fluid mixture with the collision step 
able to recover the hydrodynamics and non-ideal mixing dynamics. The free energy model utilizes phase field theory and constructs a free energy 
functional to recover interfacial dynamics and effects in a thermodynamically consistent manner. Often, a square gradient free energy is used for 
simplicity and ease of implementation. The mean field theory The non-ideal mixing dynamics are recovered during the "recoloring" step, of which the 
technique proposed by Latva-Kokko and Rothman is more commonly used for soft matter flows. \cite{liu_multiphase_2016}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale = 1]{figures/methods/d3q19_lattice.jpg}
    \caption{D3Q19 lattice demonstrating the rest, nearest and next nearest direction that correspond to the 19 
    directions $(i)$ of the lattice with lattice velocity $c_{i}$. \cite{schmieschek_lb3d_2017} Reproduced from 
    Schmieschek et al. Computer Physics Communications 2017, 217, 149--161, under the Creative Commons license.}
    \label{fig:d3q19_lattice}
\end{figure}

@@

% \textcolor{blue}{https://doi.org/10.1103/PhysRevE.84.046710}
